Transfer report

### 2018-09-21 Report/paper-story.Rmd
Updates the paper report

block_measure <- function(df, value){
  df_wide <- dcast(df, learning.condition + id ~ exp_block_id, value.var = value)
  colnames(df_wide) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
  df_wide %>% 
    group_by(learning.condition) %>%
    summarise(block1.mean  = mean(block1, na.rm = T),
              block3.mean = mean(block3, na.rm = T),
              block4.mean = mean(block4, na.rm = T),
              block34.diff = mean((block3-block4)/(block4+block3), na.rm = T),
              block34.se  = sqrt(var((block3-block4)/(block4+block3), na.rm = T)/sum(!is.na(block3-block4))),
              block14.diff = mean((block1-block4)/(block1+block4), na.rm = T),
              block14.se = sqrt(var((block1-block4)/(block1+block4), na.rm = T)/sum(!is.na(block1-block4)))
  )
}
```
 There will be smaller spatial knowledge transfer in those conditions which use less

@@

## Descriptives
In normalised distance we can see that people in different conditions start fairly at the same level of performance (difference between learnign conditions is not significant `r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`). Nor there is significant difference between all 5 starting groups `r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$condition`.

```{r, echo=F, warning=F}
make_graph(sub_walk_all, "learning.condition", "min_norm_distance", "exp_block_id") + 
  ylab("Average normalised trial distance") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal,"Block ID")

@@

We see that both real world and treadmill VR learning leads to improvement from block 1 to block 4 in both errors and path travelled. But desktop only in number of errors and marginally in distance improvement.

### Overrall average walking and error performance in phase 1 vs phase 2
Given that all participants had 3 blocks of trials in the second phase done in the real world, We can assume, if one learning method leads to better transfer than another, than the performance improvement between phases could be different.
```{r, echo=F, warning=F}
make_graph(sub_walk_all, "learning.condition", "min_norm_distance", "phase") + 
  ylab("Average normalised trial distance") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal,"Phase")
make_graph(sub_walk_all, "learning.condition", "errors", "phase")+ 
  ylab("Average number of errors") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal,"Phase")
```

Anova shows a significant effect of phase and condition on overall performance, but not of their interaction.
```{r, echo=F}
apa_print(aov(min_norm_distance ~ phase*learning.condition, sub_walk_all))
pair_t_norm_distance_phase_2 <- pairwise.t.test(sub_walk_all$min_norm_distance, sub_walk_all$learning.condition, p.adj = "holm")
```
Looking solely at the performance differences in the second phase differs significantly between groups (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$phase == 2,]))$statistic$learning_condition`). Tukey test shows significant difference between real adn a desktop versions

```{r, echo=F}
kable(TukeyHSD(x=aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$phase == 2,]), conf.level=0.95)$learning.condition[, c(1,4)])
```

### Overrall average pointing performance in phase 1 vs phase 2
```{r, echo=F, warning=F}
make_graph(sub_sop_all, "learning.condition", "abs_error", "phase") + 
  ylab("Average normalised trial distance") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal,"Phase")
```

Anova shows significant effect of both phase, learning condition and their interaction on pointing performance.
```{r, echo=F}
kable(apa_print(aov(abs_error ~ phase*learning.condition, sub_sop_all))$table)
```

@@

Looking directly at the performance after the switch, we can see if there is a sudden drop in performance. We can study this both suing individual paired t.tests as well as non paired t-tests. Looking at the graphs we see differences in the level to which people have learned the environemtn in the 3rd block.  
```{r, echo=F, message=F}
make_graph(sub_walk_all[sub_walk_all$exp_block_id %in% c(3,4),], "learning.condition", "min_norm_distance", "exp_block_id") + 
  ylab("Average normalised trial distance") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal,"Block ID")
make_graph(sub_walk_all[sub_walk_all$exp_block_id %in% c(3,4),], "learning.condition", "errors", "exp_block_id") + 
  ylab("Average number of errors") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal,"Block ID")
```

Anova doesn¨t show shows a significant difference in the block 3 (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 3,]))$statistic$learning_condition`), but there is a significant difference between path performance in the block 4 (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`).

There is significant difference in the number of errors in both block 3 (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 3,]))$statistic$learning_condition`), and the block 4 (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`).

Running pairwise t-tests to see performance change from block 3 to 4 gives the following results.
```{r, echo=F}
df_dist <- block_t_test(sub_walk_all, 3, 4,"min_norm_distance")[,c(1,4)]
df_error <- block_t_test(sub_walk_all, 3, 4,"errors")[,c(1,4)]
df34 <- merge(df_dist, df_error, by = "learning.condition")
colnames(df34) <- c("learning condition", "p-value distance improvement", "p-value errors improvement")
kable(df34, caption="Paired T test comparing infividual performance improvement from block 3-4 in different conditions")
```
WE can see that all participants improved in the errors made, but neigher group is significantly worse between block 3 and 4 in the path travelled (although participants learning on desktop perform slightly worse in 4th block, but the significance is only marginal)

### Personal walking performance change in block 3 vs 4 
(calculated as an average of individuals (block3-block4)/(block3+block4))
```{r, echo=F}

### Personal walking performance change in block 1 vs 4 
(calculated as an average of individuals (block1-block4)/(block3+block4))
```{r, echo=F}

@@

```
But I can look at either pure changes between block 3-4 or 1-4 or those calculations that you proposed as a (block4-block3)/(block3+block4) measurements.

```{r, echo=F}
df_wide_errors <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "sum.errors")
df_wide_distance <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "mean.distance")
colnames(df_wide_errors) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
colnames(df_wide_distance) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")

df_wide_errors$block14 <- (df_wide_errors$block1-df_wide_errors$block4)/(df_wide_errors$block1+df_wide_errors$block4)
df_wide_errors$block34 <- (df_wide_errors$block3-df_wide_errors$block4)/(df_wide_errors$block3+df_wide_errors$block4)
df_wide_distance$block14 <- (df_wide_distance$block1-df_wide_distance$block4)/(df_wide_distance$block1+df_wide_distance$block4)
df_wide_distance$block34 <- (df_wide_distance$block3-df_wide_distance$block4)/(df_wide_distance$block3+df_wide_distance$block4)

df_errors <- block_measure(condition_all_sums, "sum.errors")
df_distance <- block_measure(condition_all_sums, "mean.distance")

clean_walk_14 <- ez_prepare_block(sub_walk_all, c(1,4))
clean_walk_34 <- ez_prepare_block(sub_walk_all, c(3,4))
```
## Block improvement from block 1 to block 4

But anova shows no significant difference in the improvement between groups for distance (`r apa_print(aov(block14 ~ learning.condition, df_wide_distance))$statistic$learning_condition`). This might be due to quite large variability in the sample.

### Errors
Unlike for distance, anova reveals significant difference between learning conditions in number of errors improvement from block 1 to block 4 (`r apa_print(aov(block14 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).

```{r, echo=F}
kable(TukeyHSD(x=aov(block14 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition[, c(1,4)])
```

```{r, echo=F}
ggplot(df_errors, aes(x = learning.condition, y = block14.diff, fill = learning.condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), width=.2, position=position_dodge(.9)) +
  scale_fill_manual(values = estimote_pal) +
  ylab("Mean sum of errors improvement from block 1 to 4") + xlab("Learning condition")  + guides(fill=F)
```

## Block performance improvement from block 3 to block 4

Everybody's distance performance (including Real to Real condition) gets a little bit worse from 3 to 4. But this is not significant when computed using paired t-tests, although it is marginally worse in the desktop version. But all conditions show decrease in the number of errors from block 3 to 4.

```{r, echo=F}
df_dist <- block_t_test(sub_walk_all, 3, 4,"min_norm_distance")[,c(1,4)]
df_error <- block_t_test(sub_walk_all, 3, 4,"errors")[,c(1,4)]
df34 <- merge(df_dist, df_error, by = "learning.condition")
colnames(df34) <- c("learning.condition", "p-value distance improvement", "p-value errors improvement")
kable(df34)
```

### Distance

Comparing the block improvement in different conditions using anovas, we see marginally significant difference in the improvement between groups for distance (`r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition`).

Within subject anova for distance improvement between 3 and 4 shows significant interaction effect and marginally significant effect of learning condition and block on distance.
```{r, warning = F, echo=F, message=F}
aov_block_dist_34 <- ezANOVA(
  data = clean_walk_34,
  dv = min_norm_distance, 
  wid = id,
  within = exp_block_id,
  between = learning.condition, 
  type = 3)
kable(aov_block_dist_34)

### Errors

ggplot(clean_walk_34, aes(x = as.numeric(exp_block_id), y = errors, color = factor(condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average number of errors") + xlab("Experimental block") + 
  scale_color_manual(values = estimote_pal, "Learning condition")
```

>>>>

block_measure <- function(df, value){
  df_wide <- dcast(df, learning.condition + id ~ exp_block_id, value.var = value)
  colnames(df_wide) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
  df_wide %>% 
    group_by(learning.condition) %>%
    summarise(block1.mean  = mean(block1, na.rm = T),
              block3.mean = mean(block3, na.rm = T),
              block4.mean = mean(block4, na.rm = T),
              block34.diff = mean((block3-block4)/(block4+block3), na.rm = T),
              block34.se  = sqrt(var((block3-block4)/(block4+block3), na.rm = T)/sum(!is.na(block3-block4))),
              block14.diff = mean((block1-block4)/(block1+block4), na.rm = T),
              block14.se = sqrt(var((block1-block4)/(block1+block4), na.rm = T)/sum(!is.na(block1-block4)))
  )
}
df_errors <- block_measure(condition_all_sums, "sum.errors")
df_distance <- block_measure(condition_all_sums, "mean.distance")

df_wide_errors <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "sum.errors")
df_wide_distance <- dcast(condition_all_sums, learning.condition + id ~ exp_block_id, value.var = "mean.distance")
colnames(df_wide_errors) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
colnames(df_wide_distance) <- c("learning.condition","id","block1", "block2","block3","block4","block5","block6")
```

@@

## Descriptives
In normalised distance we can see that people in different conditions start fairly at the same level of performance (difference between learning conditions is not significant `r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`). Nor there is significant difference between all 5 starting groups `r apa_print(aov(min_norm_distance ~ condition, walk_all[walk_all$exp_block_id == 1, ]))$statistic$condition`.

```{r, echo=F, warning=F, message=F}
make_graph(sub_walk_all, "learning.condition", "min_norm_distance", "exp_block_id") + 
  ylab("Average normalised trial distance") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal,"Block ID")

@@

We see that both real world and treadmill VR learning leads to improvement from block 1 to block 4 in both errors and path travelled. But desktop only in number of errors and marginally in distance improvement. All this suggests transfer is happening across all learning conditions.

### Overrall average walking and error performance in phase 1 vs phase 2
Given that all participants had 3 blocks of trials in the second phase done in the real world, We can assume, if one learning method leads to better transfer than another, than the performance improvement between phases could be different as well.
```{r, echo=F, warning=F,message=F}
make_graph(sub_walk_all, "learning.condition", "min_norm_distance", "phase") + 
  ylab("Average normalised trial distance") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal,"Phase")
make_graph(sub_walk_all, "learning.condition", "errors", "phase")+ 
  ylab("Average number of errors") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal,"Phase")
```

Anova shows a significant effect of phase and condition on overall performance, but not of their interaction.
```{r, echo=F}
kable(apa_print(aov(min_norm_distance ~ phase*learning.condition, sub_walk_all))$table)
pair_t_norm_distance_phase_2 <- pairwise.t.test(sub_walk_all$min_norm_distance, sub_walk_all$learning.condition, p.adj = "holm")
```
Performance in the second phase differs significantly between groups (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$phase == 2,]))$statistic$learning_condition`), suggesting that the learnign modality has an effect on the overall performance. Tukey test shows significant difference between real adn a desktop versions. although this can be only because of bad performance in block 4 after switch, not due to blocks 5 and 6 (will test later.)

```{r, echo=F}
df_temp <- TukeyHSD(x=aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$phase == 2,]), conf.level=0.95)$learning.condition[, c(1,4)]
kable(tukey_report(df_temp))
```

### Overrall average pointing performance in phase 1 vs phase 2
The issue with comparing pointing is that people point after learning condition and then only after being subjected to the new condition for 18 trials. Which means that each participant has experienced the real world for half an hour before pointing. As we see laterr, the biggest difference in transfer seems to occur between blocks 3-4, with most participants achieving same performance by the block 6. That means that the comparison of pointing is not very informative.

```{r, echo=F, warning=F, message=F}
make_graph(sub_sop_all, "learning.condition", "abs_error", "phase") + 
  ylab("Average normalised trial distance") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal,"Phase")
```

Anova shows significant effect of both phase, learning condition and their interaction on pointing performance. Probably due to the fact that participants pointed less acurately on the treadmill and evn less on the desktop, but all then achieved same performance by the end of the 6th block.
```{r, echo=F}
kable(apa_print(aov(abs_error ~ phase*learning.condition, sub_sop_all))$table)
```

@@

Looking directly at the performance after the switch, we can see if there is a sudden drop in path length desktop condition. We can study this using paired t.tests as well as non paired t-tests. Looking at the graphs we see differences in the level to which people have learned the environemtn in the 3rd block.

```{r, echo=F, warning=F, message=F}
make_graph(sub_walk_all[sub_walk_all$exp_block_id %in% c(3,4),], "learning.condition", "min_norm_distance", "exp_block_id") + 
  ylab("Average normalised trial distance") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal,"Block ID")
make_graph(sub_walk_all[sub_walk_all$exp_block_id %in% c(3,4),], "learning.condition", "errors", "exp_block_id") + 
  ylab("Average number of errors") + xlab("condition") + 
  scale_fill_manual(values=estimote_pal,"Block ID")
```

Anova doesn't show shows a significant difference between groups in the block 3 (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 3,]))$statistic$learning_condition`), but there is a significant difference between distance performance in the block 4 (`r apa_print(aov(min_norm_distance ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`).

There is significant difference in the number of errors in both block 3 (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 3,]))$statistic$learning_condition`), and the block 4 (`r apa_print(aov(errors ~ learning.condition, sub_walk_all[sub_walk_all$exp_block_id == 4,]))$statistic$learning_condition`)., with treadmill and desktop performing worse than real world.

Running pairwise t-tests to see performance change from block 3 to 4 gives the following results.
```{r, echo=F}
df_dist <- block_t_test(sub_walk_all, 3, 4,"min_norm_distance")[,c(1,4)]
df_error <- block_t_test(sub_walk_all, 3, 4,"errors")[,c(1,4)]
df34 <- merge(df_dist, df_error, by = "learning.condition")
colnames(df34) <- c("learning condition", "p-value distance improvement", "p-value errors improvement")
kable(df34, caption="Paired T test comparing infividual performance improvement from block 3-4 in different conditions")
```

WE can see that all participants improved in the errors made, but neigher group is significantly worse between block 3 and 4 in the path travelled (although participants learning on desktop perform slightly worse in 4th block, but the significance is only marginal)

### Personal walking performance change in block 3 vs 4

```{r, echo=F}
df_wide_errors$block34 <- (df_wide_errors$block3-df_wide_errors$block4)/(df_wide_errors$block3+df_wide_errors$block4)
df_wide_distance$block34 <- (df_wide_distance$block3-df_wide_distance$block4)/(df_wide_distance$block3+df_wide_distance$block4)
clean_walk_34 <- ez_prepare_block(sub_walk_all, c(3,4))
```
The other approach is using the "performance change" coefficient as you proposed calculated as an average of individual's (block3-block4)/(block3+block4))

Comparing the block perfomance change in different conditions using anovas, we see marginally significant difference in the improvement between groups for distance (`r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition`).

Within subject anova for distance improvement between 3 and 4 shows significant interaction effect and marginally significant effect of learning condition and block on distance.

```{r, warning = F, echo=F, message=F}
aov_block_dist_34 <- ezANOVA(
  data = clean_walk_34,
  dv = min_norm_distance, 
  wid = id,
  within = exp_block_id,
  between = learning.condition, 
  type = 3)
kable(aov_block_dist_34)

ggplot(clean_walk_34, aes(x = as.numeric(exp_block_id), y = errors, color = factor(condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average number of errors") + xlab("Experimental block") + 
  scale_color_manual(values = estimote_pal, "Learning condition")
```
### Personal walking performance change in block 1 vs 4 
(calculated as an average of individuals (block1-block4)/(block3+block4))
```{r, echo=F}
df_wide_errors$block14 <- (df_wide_errors$block1-df_wide_errors$block4)/(df_wide_errors$block1+df_wide_errors$block4)
df_wide_distance$block14 <- (df_wide_distance$block1-df_wide_distance$block4)/(df_wide_distance$block1+df_wide_distance$block4)
clean_walk_14 <- ez_prepare_block(sub_walk_all, c(1,4))
```

Anova shows no significant difference in the improvement between groups for distance (`r apa_print(aov(block14 ~ learning.condition, df_wide_distance))$statistic$learning_condition`). This might be due to quite large variability in the sample.

But it reveals significant difference between learning conditions in number of errors improvement from block 1 to block 4 (`r apa_print(aov(block14 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).

```{r, echo=F}
kable(TukeyHSD(x=aov(block14 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition[, c(1,4)])
```
```{r, echo=F}
ggplot(df_errors, aes(x = learning.condition, y = block14.diff, fill = learning.condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), width=.2, position=position_dodge(.9)) +
  scale_fill_manual(values = estimote_pal) +
  ylab("Mean sum of errors improvement from block 1 to 4") + xlab("Learning condition")  + guides(fill=F)
```

@@
